---
title: "Homework 3"
output: github_document
---

Starting assignment and loading data set:

```{r, message=FALSE}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

# Problem 1:

*Description:*
The dataset for Instacart contains `r ncol(instacart)` variables and `r nrow(instacart)` observations. Some key variables include identifiers such as `product_id`, `user_id` and `order_id`. Other descriptory variables include `product_name`, `department`, `reordered` and `order_dow`.  The mean number of days between orders is `r instacart %>% pull(days_since_prior_order) %>% mean() %>% round(digits = 1)` days and the most popular department for orders is *`r instacart %>% pull(department) %>% max()`*. The most ordered item has been *`r instacart %>% count(product_name, sort = TRUE) %>% slice(1) %>% pull(product_name)`*, 


while the least ordered item is *`r instacart %>% count(product_name) %>% arrange(n, descending) %>% slice(1) %>% pull(product_name)`*.





### How many aisles are there, and which aisles are the most items ordered from?
There are a total of `r instacart %>% pull(aisle_id) %>% n_distinct()` aisles, with the `r instacart %>% count(aisle, sort = TRUE) %>% slice(1) %>% pull(aisle)` aisle being the one with the most items ordered, 

### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r, message=FALSE}
instacart %>%
  count(aisle, sort = TRUE) %>%
  filter(n > 10000) %>%
  ggplot(aes(x = reorder(aisle,n), y = n)) + 
    geom_col() +
    coord_flip() +
    labs(x = "Aisle", y = "Count of Items Ordered ", title = "Aisles by count of items ordered") +
    geom_text(aes(label = n), hjust = -0.1, colour = "purple", size = 3) +
    scale_y_continuous(limits = c(-1, 165500)) +
    theme(plot.title = element_text(hjust = 0.1), 
        axis.title.x = element_text(face = "bold", size = 12),
        axis.title.y = element_text(face = "bold", size = 12))
```

### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r, message=FALSE}
instacart %>%
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>%
  summarise(times_ordered = n()) %>%
  arrange(aisle, desc(times_ordered)) %>%
  slice(1:3) %>%
  knitr::kable()
```

### Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r, message=FALSE}
apple_coffee = instacart %>%
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  select(product_name, order_dow, order_hour_of_day) %>%
  arrange(order_dow) 

apple_coffee %>%
  pivot_wider(
    names_from = "product_name",
    values_from = "order_hour_of_day",
    values_fn = mean
  ) %>%
  knitr::kable(digits = 0)
```

